\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements} % Add the preface to the table of contents as a chapter

This thesis would not have been possible without the support and work of all those who surrounded me, both
professionally and personally.

First of all, I am deeply grateful to my supervisor, Imma Riu, for her endless support, invaluable scientific advise and perpetual motivation to grow as a physicist. I appreciate all her time and opportunities given since my first contact with particle physics as an undergraduate student. I would also thank Lluïsa Maria Mir, Martine Bosman and Aurelio Juste for their work, guidance and contagious enthusiasm for this field all these years. I am very grateful for all my colleagues at IFAE and CERN that have furthermore enriched my work experience or simply shared moments with me.

Many thanks to Julian Glatzer for his essential initial guide at the start of the PhD, together with Nicola Orlando, always available despite his many activities. Also Sergio Gonzalez, with whom I started sharing office before the PhD and Carlos Moreno, the "senior" PhD of the office; despite being IFAE colleagues in different activities, we have shared a lot of the path, made fun memories and kept each other sane. I also have to thank all the colleagues from the charged Higgs and HBSM groups in ATLAS for all the
important advises to the developments of the data analyses. I am also thankful to the ATLAS tau trigger group and their coordinators, especially Bertrand and Antonio, who always quickly answer my questions with invaluable insight on the taus and Athena.

Being at CERN has been absolutely fantastic and one of the best environments to work in. Not only because it is the center of experimental particle physics, but the amazing people from all over the world that contribute to its atmosphere. I will always be grateful for the friends and great people I have met through work or drinks -- Víctor, César, Chiara, Tal, José L., Trisha, Marton, Matthijs, Antoine, Max, Péter, Stergios, Anil, Tim, Lisa, Eyleen, Martí, Sabino, Francesca, Razvan, Florian, Mihi and Roman.

I would also like to thank students that have joined along the way some of which I have had the pleasure of welcoming, tried to motivate or taught a thing or two: Adrià B., Shalini, Paula, Shubham, Simon and Jack.

I need to mention people outside the field for their great support and that have been there in the good and bad times: Blanca and Chris, Sergio, Helena, Matt, Andreu, Julián, Neus and Dende with his parents.

Finalment, gràcies Xavi per ser allà des de que tinc memòria i per ser un increïble exemple de constància i esforç. Per últim, però la més important, la meva font principal de felicitat, motivació i suport malgrat la distància; es gràcies a tú, Marta, que he arribat on sóc; tot ha anat segons el pla.